Response to the reviewers’ comments on the article

### 2020-04-03 article_springer/article_springer.tex
New proof inserted and points of Peter (email 29/03/2020) adressed.

\emph{Proof of Theorem:} We interpret the threshold crossing time
\(T(\ell)\) as the hitting time of the underlying CTRM (Continuous Time
Random Maxima) or ``max-renewal process'', and then utilize a result by
Meerschaert and Stoev (2008). The running maximum process is defined as
\[
M(c) := J_1 \vee \ldots \vee J_{\lfloor c \rfloor},
\] and since we assumed that the \(J_k\) our out of the domain of
attraction of a max-stable distribution \(F\) (see
\eqref{assumptionJs}), there exist norming functions \(a(c)\) and
\(d(c)\) such that \[
\mathbf P\left[ \frac{M(c) - d(c)}{a(c)} \le \ell^* \right] 
\longrightarrow F(\ell^*), \quad c \to \infty
\] where \(F\) is a generalized extreme value distribution, and
\(\ell^*\) is any value from the support of \(F\). The CTRM process is
then defined via \[
V(t) = M(N(t)), \quad t \ge 0
\] where \(N(t)\) is the renewal process associated with the waiting
times \(W_k\): \[
N(t) = \max\{n: W_1 + \ldots + W_n \le t\}.
\] Now a key observation is that \[
T(\ell) = \inf\{t: V(t) > \ell\}, 
\] and that

\begin{align}
T(\ell) > t \quad \text{ if and only if } \quad V(t) \le \ell. \label{inverserelat}
\end{align}

By (Theorem 3.1, Meerschaert and Stoev, 2008), we have the stochastic
process convergence

\begin{align}
\frac{V(ct) - d(\tilde b(c))}{a(\tilde b(c))} 
\stackrel{d}{\longrightarrow} Y(t),  \quad c \to \infty, \label{TheoMS}
\end{align}

for \(t > 0\), where \(Y(t)\) is a time-changed (``subordinated'')
extremal process, and where \(\tilde b(c)\) is a regularly varying
norming function which is \emph{inverse} to \(b(c)\), in the sense that
\(b(\tilde b(c)) \sim c \sim \tilde b(b(c))\) (Bingham et al., 1989).

Without loss of generality, we choose \(\ell^*\) such that
\(F(\ell^*) = 1/e\), and let
\(\ell = a(\tilde b(c)) \ell^* + d(\tilde b(c))\). We may then calculate
\[
\mathbf P\left[ \frac{T(\ell)}{b(1/p_{\ell})} > t \right]
= \mathbf P[T(\ell) > b(1/p_{\ell}) t]
= \mathbf P[V(ct) \le \ell]
\] where we have substituted \(c = b(1/p_{\ell})\) and used
\eqref{inverserelat}. Moreover, we receive due to the convergence in
\eqref{TheoMS} \[
\mathbf P[V(ct) \le \ell]
= \mathbf P\left[ \frac{V(ct) - d(\tilde b(c))}{a(\tilde b(c))} 
\le \frac{\ell - d(\tilde b(c))}{a(\tilde b(c))} \right]
\longrightarrow \mathbf P\left[ Y(t) \le \ell^* \right] ,  \quad c \to \infty,
\] for \(t>0\). Defining the hitting time of level \(\ell^*\) by
\(Y(t)\) as \(\xi_{\ell^*} := \inf\{t: Y(t) > \ell^*\}\), we then have
\[
P\left[ Y(t) \le \ell^* \right] = \mathbf P[\xi_{\ell^*} > t] 
= \mathbf P[(-\log F(\ell^*))^{-1/\beta} X^{1/\beta} D > t]
\] by (Proposition 4.2, Meerschaert and Stoev, 2008), where \(X\) is an
exponential random variable with mean \(1\). Using (Theorem 19.1,
Haubold et al., 2011), we see that
\(X^{1/\beta} D \sim {\rm ML}(\beta, 1)\), concluding the proof. \qed

@@

The classical POT approach is based on the fact that exceedances above a
high threshold are asymptotically GPD distributed, hence a GPD
distribution is fitted to the exceedances above a high enough threshold.
This approach was well-established a long time ago and is standard in
extreme value theory nowadays. Hence we just refer here to the according
literature (Smith (1984), Leadbetter (1991), Davison and Smith (1990)),
a nice introduction can be found in Beirlant et al. (2006) Ch.5.3,
Embrechts et al. (2013) Ch. 6.5. or Coles (2001) Ch.4. Hence, we focus
in this article mainly on the inter-exceedance times. Due to the Theorem
in the last Section we know that for high thresholds the times between
exceedances are asymptotically Mittag-Leffler distributed in case of
heavy-tailed waiting times between the observations. In this article we
propose to proceed with inter-exceedance times in a similar fashion as
with the magnitudes in the classical POT approach, more precisely to fit
a Mittag-Leffler distribution via stability plots to the
inter-exceedance times. Hence, before we talk about inference for the
inter-exceedance times in the next section, we first discuss inference
for Mittag-Leffler distributions.

@@

Since the Mittag-Leffler distribution is heavy-tailed, many researchers
would intuitively give the highest importance to the tail behaviour of
the distribution. Of course, one can also use established tail exponent
estimators for the estimation of the parameter \(\beta\) such as the
Hill estimator (Hill, 1975) based on the \(r+1\) upper order statistics
\begin{align}\label{eq:Hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align} where \(X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}\)
denote the order statistics in decreasing order of a sequence
\(X_1,\ldots,X_n\). The Hill estimator only performs reliably well for
distributions close to Pareto (see Resnick, 1997). It heavily depends on
the choice of \(r\) in \eqref{eq:Hill}, hence on the number of order
statistics entering into \eqref{eq:Hill}. Figure \ref{fig:Hillplots}
shows Hill plots for Mittag-Leffler simulated data, with varying sample
sizes and tails. To deduce the correct tail parameter estimates from
these plots is virtually impossible. To estimate the tail parameter of a
Mittag-Leffler distribution with the Hill estimator, the sample size has
to be much larger and one has to choose a large tuning parameter \(r\).
In case that only events above a high threshold are recorded,
\(n=10000\) events are unrealistic. There is a vast literature on
bias-reduced versions of the Hill estimator and on the optimal choice of
\(r\), mainly based on imposed second-order regular variation conditions
(see e.g.~Beirlant et al., 2006, p. Ch.4.4 and 4.7 and references in
it). Also, there are lot of other tail-index estimators not taken under
consideration in our article, since this would go beyond the scope of
our article. However, we know from the Theorem in Section
\ref{sec:scaling} that the inter-exceedance times are asymptotically
Mittag-Leffler distributed. So all these general tail (and scale)
estimation methods are statistically less efficient since they only use
a portion of the information contained in the data, as mentioned by
Kozubowski (2001). Furthermore, the Hill estimator of course completely
fails, if the inter-arrival times are exponentially distributed and not
heavy-tailed. Hence, we propose to use the log-Moment estimator or the
MLE for Mittag-Leffler distributions.

@@

{\centering \includegraphics[width=0.9\linewidth]{article_springer_files/figure-latex/Hillplots-1}

@@

\caption{\label{fig:Hillplots} Hillplots for simulated Mittag-Leffler data with true tail 0.8 and sample size 200 (left) and 10000 (right), with number of upper order statics r on which the Hill estimator is based on the x-axis. }\label{fig:Hillplots}
\end{figure}

@@

  \begin{enumerate}
  \def\labelenumii{\alph{enumii})}
  \tightlist
  \item
    extract the set \(\mathcal T_k\) of exceedance times between the
    magnitudes exceeding the threshold \(X_{(k)}\)
  \item
    fit a Mittag-Leffler distribution to \(\mathcal T_k\), resulting in
    the parameter estimates \(\hat\beta_k\) and \(\hat \sigma_k\).
  \end{enumerate}
\item
  Plot \(k\) vs.~\(\hat \beta_k\). To the right, as
  \(k \downarrow K_0\), the asymptotics are off and bias is high. To the
  left (high threshold), data are scarce and variance is high. Choose a
  region of \emph{stability} in the middle for a parameter estimate
  \(\hat \beta\).
\item
  Plot \(k\) vs.~\(k^{1/\hat \beta} \hat \sigma_k\). Again, choose a
  region of \emph{stability} as \(k\) gets smaller and define
  \(\hat \sigma_0\).
\end{enumerate}

@@

The performance of log-Moment and Maximum-Likelihood estimators for the
tail resp. the scale parameter is shown in Figure \ref{Fig:TailSimu}
resp. \ref{Fig:ScaleSimu}. Both estimators show good performance, with a
slight advantage for the Maximum-Likelihood estimator. This advantage is
paid for by a higher computational cost. Moreover, the bottom row in
Figure \ref{Fig:TailSimu} shows clearly that the log-Moment and
Maximum-Likelihood estimators generalize to the exponential case
(\(\beta = 1\)) whereas a general tail estimator like the Hill-estimator
would not.

@@

The bursty CTRE approach to model ``non-Poissonian'' threshold crossing
times should be contrasted with the (now standard) approach of clusters
of extremes, see e.g.~Ferro and Segers (2003). In this approach, i.i.d.
event sequences of magnitudes are generalized to stationary sequences of
event magnitudes (subject to a mixing condition). The two approaches are
fundamentally different: A clustering model assumes that each event
belongs to one particular (latent) group of events. For bursts, however,
the aim is to identify an underlying scale-free pattern in the event
dynamics, which is often characteristic of complex systems. It is an
interesting open problem to develop quality criteria, based e.g.~on
measures of surprise (Lee et al., 2015), which guide an applied
statistician in the choice between a clustering and a CTRE approach for
a particular problem. But the CTRE approach should not be seen as a
substitute for the usual cluster approach. It was for example also
observed in Towe et al. (2020) that clusters of extreme events still
exist after declustering. We believe it is possible to unify the cluster
approach with our CTRE approach by considering CTREs based on MRPs with
a \emph{stationary}, rather than i.i.d., sequence of magnitudes. Then
the times between clusters are asymptotically Mittag-Leffler
distributed. The identification of clusters in the context of
heavy-tailed inter-arrival times is an open problem. But what one could
do in case of observing dependencies between the exceedances: one could
build natural clusters for example by considering daily maxima instead
of hourly maxima.

Finally, a purely scale-free pattern for event times may be too rigid as
an assumption for some bursty time series, because often the
heavy-tailed character of the inter-arrival time distribution does not
hold at all time scales; rather, it applies at short and intermediate
time scales, and is truncated (or tempered, reverting to an exponential
distribution) at very long time scales (see e.g.~Meerschaert et al.,
2012; and Aban et al., 2006). In such situations, a ``tempered''
Mittag-Leffler distribution may provide a more realistic fit, which we
aim to introduce in follow-up work.

@@

\leavevmode\hypertarget{ref-BinghamGoldieTeugels}{}%
Bingham, N.H., Goldie, C.M., Teugels, J.L., 1989. Regular variation.
Cambridge university press.

@@

@@

@@

@@

@@

>>>>

\emph{Proof of Theorem:} Due to assumption \eqref{assumptionJs}, it
follows \begin{align*}
    P\left(\frac{M(c)-d(c)}{a(c)} \leq x\right) \rightarrow F(x) \text{ as } c \rightarrow \infty,
    \end{align*} where \(x\) is any value from the support of \(F\).
Then it follows, \begin{align*}
      F(xa(c)+d(c))^{\lfloor c \rfloor} &\rightarrow F(x) \text{ as } c \rightarrow \infty,\\
      \Rightarrow  \lfloor c \rfloor \log F(\ell_c)  &\rightarrow \log F(x)\text{ as } c \rightarrow \infty,
    \end{align*} with \(\ell_c:=xa(c)+d(c)\). Furthermore, since
\(\log(1-x) \sim -x\) for small \(x\) it follows \begin{align}
      c \cdot p(\ell_c) \rightarrow -\log F_A(x)  \text{ as } c \rightarrow \infty,\label{eq:lplconv}
    \end{align} with \(p(\ell_c):= \overline{F}(\ell_c)\). Since
\(\tau(\ell_c) \sim Geo(p(\ell_c))\) and \eqref{eq:lplconv} it follows
that \(\tau(\ell_c)/c\) converges to an Exponential distributed random
variable, \begin{align*}
    \frac{\tau(\ell_c)}{c} \Rightarrow E  \text{ as } c \rightarrow \infty
    \end{align*} with parameter \(\lambda:=-\log F_A(x)\). Due to
\begin{align*}
    S(c):= \sum_{i=1}^{\lfloor c \rfloor} \frac{W_i}{b(c)} \Rightarrow D \ \text{ as } c \rightarrow \infty,
    \end{align*} it follows with Gnedenkos transfer theorem (see
Gnedenko (1983)), that \begin{align}
    \sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(\ell_c)}  \Rightarrow Z\ \text{ as }  c \rightarrow \infty, \label{TheoGutArg}
    \end{align} where the distribution of \(Z\) has the characteristic
function \begin{align}
    \Psi(t) &= \int_0^{\infty} (\Phi_D(t))^{y} F_E(dy) \nonumber\\
    &= \int_0^{\infty} \frac{1}{1-\log(\Phi_D(t))/\lambda} \nonumber\\
    &= \frac{1}{1+s^{\beta}\lambda^{-1}}
    \end{align} where \(\Phi_D\) resp. \(F_E\) denotes Laplace transform
of \(D\) resp. the distribution function of \(E\). Hence, \(Z\) is
Mittag-Leffler distributed with parameter \(\lambda^{-1/\beta}\). Since
\begin{align*}
    \sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(1/p(\ell_c))} = \left(\sum_{i=1}^{ \tau(\ell_c)} \frac{W_i}{b(c) }\right) \frac{b(c)}{b(1/p(\ell_c))}  
    \end{align*} and due to \begin{align*}
    \frac{b(c)}{b(1/p(\ell_c))} = \frac{b(c)}{b(c/(p(\ell_c)c))} \sim (p(\ell_c) \cdot c)^{1/\beta} \rightarrow \lambda^{1/\beta} \text{ as } c \rightarrow \infty.
    \end{align*} the second factor converges to \(\lambda^{1/\beta}\),
it follows that \(Z \sim ML(\beta,1)\). Since \(c \rightarrow \infty\)
is equivalent to \(\ell \uparrow x_R\) the assertion follows with
\(\ell:=\ell_c\) and \(p_\ell:=p(\ell_c)\). \qed

@@

The classical POT approach is based on the fact that exceedances above a
high threshold are asymptotically GPD distributed, hence a GPD
distribution is fitted to the exceedances above a high enough threshold.
As this approach is by now well-established in extreme value theory, we
do not discuss it any further. Instead, this article focuses on our new
contribution: the modelling of inter-exceedance times. For literature
pointing to POT see e.g.~Smith (1984); Leadbetter (1991); Davison and
Smith (1990); Beirlant et al. (2006) Ch.5.3; Embrechts et al. (2013) Ch.
6.5.; Coles (2001) Ch.4.

Due to the Theorem in the last Section we know that for high thresholds
the times between exceedances are asymptotically Mittag-Leffler
distributed in case of heavy-tailed waiting times between the
observations. In this article we propose to model inter-exceedance times
in a similar fashion as magnitudes in the classical POT approach, more
precisely to fit a Mittag-Leffler distribution to the inter-exceedance
times via stability plots. Hence, before we discuss inference for the
inter-exceedance times in the next section, we first discuss inference
for Mittag-Leffler distributions.

@@

Since the Mittag-Leffler distribution is heavy-tailed, many researchers
would intuitively give the highest importance to the tail behaviour of
the distribution. Of course, one can also use established tail exponent
estimators for the estimation of the parameter \(\beta\) such as the
Hill estimator (Hill, 1975) based on the \(r+1\) upper order statistics
\begin{align}\label{eq:Hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align} where \(X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}\)
denote the order statistics in decreasing order of a sequence
\(X_1,\ldots,X_n\). The Hill estimator only performs reliably well for
distributions close to Pareto (see Resnick, 1997). It heavily depends on
the choice of \(r\) in \eqref{eq:Hill}, and hence on the number of order
statistics entering \eqref{eq:Hill}. Figure \ref{fig:Hillplots} shows
Hill plots for simulated Mittag-Leffler data, with varying sample sizes
and tails. To deduce the correct tail parameter estimates from one of
those Hillplots is virtually impossible. To estimate the tail parameter
of a Mittag-Leffler distribution with the Hill estimator, the sample
size has to be much larger and one has to choose a small tuning
parameter \(r\). In case that only events above a high threshold are
recorded, \(n=10000\) events are extremely unrealistic. There is a vast
literature on bias-reduced versions of the Hill estimator and on the
optimal choice of \(r\), mainly based on imposed second-order regular
variation conditions (see e.g.~Beirlant et al., 2006, p. Ch.4.4 and 4.7
and references in it). Also, there are hundreds other tail-index
estimators not considered in this article, since this would go beyond
the scope of our article. For example the moment estimator (Dekkers et
al., 1989), the Pickands estimator (Pickands III and others, 1975), the
QQ-estimator (Kratz and Resnick, 1996), to name a few. A nice overview
can be found in (Fedotenkov, 2018). However, we know from the Theorem in
Section \ref{sec:scaling} that the inter-exceedance times are
asymptotically Mittag-Leffler distributed, and hence it is unsurprising
that general tail (and scale) estimation methods that do not take into
account this structure are statistically less efficient (Kozubowski,
2001). Finally, the Hill estimator completely fails in the important
special case where the inter-arrival times are exponentially distributed
and not heavy-tailed. For these reasons, we propose to use the
log-Moment estimator or the MLE for Mittag-Leffler distributions.

@@

{\centering \includegraphics[width=0.5\linewidth]{article_springer_files/figure-latex/Hillplots-1}

@@

\caption{\label{fig:Hillplots} Hillplots for m=100 simulated Mittag-Leffler datasets with true tail 0.8 and sample size 200 (left) and 10000 (right), with number of upper order statistics r on which the Hill estimator is based on the x-axis. The grey thin lines are the Hill plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter. }\label{fig:Hillplots}
\end{figure}

@@

  \begin{enumerate}
  \def\labelenumii{\alph{enumii})}
  \tightlist
  \item
    extract the set \(\mathcal T_k\) of exceedance times between the
    magnitudes exceeding the threshold \(X_{(k)}\)
  \item
    fit a Mittag-Leffler distribution to \(\mathcal T_k\), resulting in
    the parameter estimates \(\hat\beta_k\) and \(\hat \sigma_k\).
  \end{enumerate}
\item
  Plot \(k\) vs.~\(\hat \beta_k\). To the right (\(k \uparrow K\), low
  threshold), the asymptotics are off and bias is high. To the left
  (\(k \downarrow K_0\), high threshold), data is scarce and variance is
  high. In the middle, look for a region of \emph{stability} for a
  parameter estimate \(\hat \beta\). Choose as \(\hat \beta\) a
  representative value from this region.
\item
  Plot \(k\) vs.~\(k^{1/\hat \beta} \hat \sigma_k\). Again, choose a
  region of \emph{stability} and, and choose a representative
  \(\hat \sigma_0\) from that region.
\end{enumerate}

@@

The performance of log-Moment and Maximum-Likelihood estimators for the
tail resp. the scale parameter is shown in Figure \ref{Fig:TailSimu}
resp. \ref{Fig:ScaleSimu}. Both estimators show good performance, with a
slight advantage for the Maximum-Likelihood estimator. This advantage is
paid for with a higher computational cost. Moreover, the bottom row in
Figure \ref{Fig:TailSimu} shows clearly that the log-Moment and
Maximum-Likelihood estimators generalize to the exponential case
(\(\beta = 1\)) whereas the Hill-estimator would not.

@@

The bursty CTRE approach to model ``non-Poissonian'' threshold crossing
times should be contrasted with the (now standard) approach of clusters
of extremes, see e.g.~Ferro and Segers (2003). In this approach, i.i.d.
event sequences of magnitudes are generalized to stationary sequences of
event magnitudes (subject to a mixing condition). The two approaches are
fundamentally different: Clustering assumes that each event belongs to
one particular (latent) group of events. For bursts, however, the aim is
to identify an underlying scale-free pattern in the event dynamics,
which is often characteristic of complex systems. It is an interesting
open problem to develop criteria, based e.g.~on measures of surprise
(Lee et al., 2015), that guide an applied statistician in the choice
between a clustering and a CTRE approach for a particular problem. In
sum, the CTRE approach should not be seen as a substitute for the usual
cluster approach, but rather as a complementary approach.

Moreover, it was observed in Towe et al. (2020) that clusters of extreme
events still exist after declustering. We believe it is possible to
marry the clustering approach with our CTRE approach by considering
CTREs based on MRPs with a \emph{stationary}, rather than i.i.d.,
sequence of magnitudes, so that the times between clusters are
asymptotically Mittag-Leffler distributed. More broadly speaking, the
identification of clusters in the context of heavy-tailed inter-arrival
times is an open problem.

Finally, a purely scale-free pattern for event times may be too rigid an
assumption for some bursty time series, because often the heavy-tailed
character of the inter-arrival time distribution does not hold at all
time scales. Rather, it applies at short and intermediate time scales,
and is truncated (or tempered, reverting to an exponential distribution)
at very long time scales (see e.g.~Meerschaert et al., 2012; and Aban et
al., 2006). In such situations, a ``tempered'' Mittag-Leffler
distribution may provide a more realistic fit, which we aim to introduce
in follow-up work.

@@

@@

\leavevmode\hypertarget{ref-dekkers1989moment}{}%
Dekkers, A.L., Einmahl, J.H., De Haan, L., 1989. A moment estimator for
the index of an extreme-value distribution. The Annals of Statistics
1833--1855.

@@

\leavevmode\hypertarget{ref-fedotenkov2018review}{}%
Fedotenkov, I., 2018. A review of more than one hundred pareto-tail
index estimators.

@@

\leavevmode\hypertarget{ref-gnedenko1983limit}{}%
Gnedenko, B., 1983. On limit theorems for a random number of random
variables, in: Probability Theory and Mathematical Statistics. Springer,
pp. 167--176.

@@

\leavevmode\hypertarget{ref-Kratz96}{}%
Kratz, M., Resnick, S.I., 1996. The qq-estimator and heavy tails.
Commun. Stat. Part C Stoch. Model. 12, 699--724.
doi:\href{https://doi.org/10.1080/15326349608807407}{10.1080/15326349608807407}

@@

\leavevmode\hypertarget{ref-pickands1975statistical}{}%
Pickands III, J., others, 1975. Statistical inference using extreme
order statistics. the Annals of Statistics 3, 119--131.

### 2020-04-03 article_springer/changes.Rmd
New proof inserted and points of Peter (email 29/03/2020) adressed.

* Thank you. We added an assumption accordingly.

@@

* We added some more explanation to the last two lines.

>>>>

* Thank you. We replaced the proof by a new one as already mentioned above. Additionaly, we added the assumption that the $J_k$ belong to some max domain of attraction at the end of Section 2.

@@

* Again this point no longer applies since we gave a completely new proof.
